
---
title: "Regression Models Course Project"
author: "Ivan Astafiev"
date: "April 3, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

Motor Trend dataset was explored to understand a relationship between transmission type (automatic/manual)and fuel usage in MilesPerGalon (MPG). The main questions was to distinguish which transmission type is more fuel efficient and try to calcuate this difference.
 
*Questions*:
 
- Is an automatic or manual transmission better for MPG
- MPG difference between automatic and manual transmissions

## Executive Summary




## DataSet exploration  

*Data Set*: mtcars
Detailed data description presented at Annex 1.

We interested in relationship of **mpg** (Miles Per Gallon) and **am** (Transmission (0 = automatic, 1 = manual). 

We should include possible confounding variables to adjust independent relationship between **mpg** and **am** variables 

```{r, include=FALSE}
library(UsingR)
library(ggplot2)
```

#### MPG to Transmission (0 = automatic, 1 = manual) 
```{r fig.width=4,fig.height=2, echo=FALSE}
g<-ggplot(data = mtcars, aes(y = mpg, x = as.factor(am), fill = as.factor(am)))+geom_violin(colour = "black", size = 2) + xlab("Transmission type") + ylab("MPG")
g
```

*MPG to Transmission* plot shows some evidence that dependency of mpg to transmission can be present. 
So, our goal is to clear that dependency from possible confounding variables to prove or reject this dependency.

## Modeling

Multivariable Regression Model is selected as a method to answer research questions.

Following process selected:

 - Create a several models with step by step adding variables
 - Compare models using Anova function to detect valuable variables
 - Constract final model with the most dependent variables
 
**Base model**
```{r}
fit0<-lm(mpg ~ am,mtcars)
```
 
**Model Selection** 
```{r, echo=FALSE }
fit1<-update(fit0,mpg ~ am + cyl,mtcars)
fit2<-update(fit0,mpg ~ am + cyl+disp,mtcars)
fit3<-update(fit0,mpg ~ am + cyl+disp+hp,mtcars)
fit4<-update(fit0,mpg ~ am + cyl+disp+hp+wt,mtcars)
fit5<-update(fit0,mpg ~ am + cyl+disp+hp+wt+qsec,mtcars)
fit6<-update(fit0,mpg ~ am + cyl+disp+hp+wt+qsec+as.factor(vs),mtcars)
fit7<-update(fit0,mpg ~ am + cyl+disp+hp+wt+qsec+as.factor(vs)+as.factor(gear),mtcars)
fit8<-update(fit0,mpg ~ am + cyl+disp+hp+wt+qsec+as.factor(vs)+as.factor(gear)+as.factor(carb),mtcars)

anova(fit0,fit1,fit2,fit3,fit4,fit5,fit6,fit7,fit8)
```

Anova results show that according to P value the following variables should be included in final model because of confirmed statistical inference: 


| Variable | Type   | Description       |
|:---------|:-------|:------------------|
|*wt*      |numeric | Weight (1000 lbs) |
|*cyl*     |factor  | Number of cylinders (to be included as Factor of 4,6,8 cylinder) |
|*hp*|numeric |Gross horsepower |

 
#### Final Model

**Model**
```{r}
final_fit<-lm(mpg~am+as.factor(cyl)+wt+hp,mtcars)
```

Adjusted R-squared:  *0.8401*

Adjusted R-squared achieved by FinalModel is a maximum value comparing to all other models tried. 

**Residuals**

```{r, echo=FALSE}
plot(final_fit,which=1)
```

Residuals spread on a plot in general without any clear dependency. However there are some dependency in a line which can be caused by outfitted values of Toyota Corolla and Fiat 128. This must be further investigated


## Executive Summary

**Final Model**

| Variable | Type   | Description       |
|:---------|:-------|:------------------|
|*am*      |factor | Transmission type (0=automatic, 1 = manual) |
|*wt*      |numeric | Weight (1000 lbs) |
|*cyl*     |factor  | Number of cylinders (to be included as Factor of 4,6,8 cylinder) |
|*hp*|numeric |Gross horsepower |

```{r}
lm(mpg~am+as.factor(cyl)+wt+hp,mtcars)
```

**Prediction precision**

Adjusted R-squared:  *0.8401*

**Coefficients**
```{r}
round(summary(final_fit)$coef,2)
```

- Increase of weight for 1000 lb cause a 2.5 miles decrease in distance driven per 1 gallon
- Usage of Manual Transmission cause a 1.8 miles increase in distance driver per 1 gallon
- Usage of 6cyl engine cause a 3 miles decrease in distance driven per 1 gallon 
- Usage of 8cyl engine cause a 2.16 miles decrease in distance driven per 1 gallon 
- P value for Transmission type do not allow to reject null hypothesis(H0), so we cant guarantee that automatic/manual transmission has a statistical dependable influence on fuel usage

**Research questions**

**Question 1**
- Is an automatic or manual transmission better for MPG?

*ANSWER* = Manual transmission is better for MPG

**Question 2**
- MPG difference between automatic and manual transmissions

*ANSWER* = Manual transmission brings 1.8 miles increase in distance driven per 1 gallon (MPG)

**However**, P value do not allow to reject null hypothesis that type of transmission has a statistical valuable influence on a MPG factor. More data and deeper research is necessary.   


## Anex 1. MTCARS DATA Descritpion


**Description**

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

**Format**

A data frame with 32 observations on 11 variables.

-	mpg	Miles/(US) gallon
-	cyl	Number of cylinders
-	disp	Displacement (cu.in.)
-	hp	Gross horsepower
-	drat	Rear axle ratio
-	wt	Weight (1000 lbs)
-	qsec	1/4 mile time
- vs	V/S
-	am	Transmission (0 = automatic, 1 = manual)
-	gear	Number of forward gears
-	carb	Number of carburetors

